Stochastic Convolutional Sparse Coding

### 2018-05-15 Stochastic Convolutional Sparse Coding/CSCmodel.tex
update

\section{Convolutional Sparse Coding (CSC)}
Specifically, the CSC problem is trying to minimize
\begin{equation}
\begin{split}
    \minimize{\filter,\code} & \frac{1}{2}\|\signal - \sum_{k=1}^{K} \filter_k * \code_k \|_2^2 + \lambda \sum_{k=1}^{K}\| \code_k \|_1 \\
    \subjectto & ~ \|\filter_k\|^2_2 \leq 1 ~~ \forall k \in {1,...K},
\end{split}
\end{equation}
where $\signal \in \mathbb{R}^D$ is a D-dimensional signal or a vectorized image~\footnote{In this manuscript, we work on 2D images.}, $\filter_k \in \mathbb{R}^M$ is the $k$-th dictionary and $\code_k$ is the sparse code associated with that dictionary, and it has the same dimension size as $\signal$. $K$ is the number of dictionary filters, and $*$ is the convolution operator. The above equations will be applied to all the training images.

The modern approaches exploit Parseval's theorem and introduce two slack variables to separate the non-smooth $L_1$ penalty term and the $L_2$ constraints, making it feasible to be efficiently computed in frequency domain, and to apply splitting strategy to formulate the problem into ADMM framework, where the two subproblems (updating $\code$ and updating $\filter$) are jointly solved by coordinate descent~\cite{bristow2013fast,heide2015fast,wohlberg2016efficient}. There exists several common issues regarding to the model and the frequency solvers.

\begin{itemize}
  \item Though CSC overcomes the independence assumption held in patch-based learning algorithms, far more variables ($K$ times more), as a payback, will be introduced to represent a single image. This brings in more severe memory and computational burden.
  \item The reconstructed sparse codes show that in the case of $K=100$, the majority of the entries (more than $99.5\%$) do not provide valid information to the represented image, which indicates the subproblem for updating $\code$ solves a highly sparse LASSO problem. Transforming the problem into frequency domain imposes restriction on the use of optimization strategies for this very specific problem.
  \item The prior work shows its efficiency of solving the CSC problem in frequency domain, while this is only applicable for updating $\code$, and it does not hold for updating $\filter$. The dictionary filters usually have much smaller spatial support than the dimension size of the sparse codes ($M \ll D$), while in order to tackle the problem in frequency domain, it requires to process the $\filter$-subproblem in the same dimensions as the sparse codes, and then project the results onto its small spatial support.
\end{itemize}

>>>>

\section{Convolutional Sparse Coding (CSC)}
Specifically, the CSC problem has the form
\begin{equation}
\begin{split}
    \minimize{\filter,\code} & \frac{1}{2}\|\signal - \sum_{k=1}^{K} \filter_k * \code_k \|_2^2 + \lambda \sum_{k=1}^{K}\| \code_k \|_1 \\
    \text{subjec to} & ~ \|\filter_k\|^2_2 \leq 1 ~~ \forall k \in \{1,\dots,K\},
\end{split}
\end{equation}
where $\signal \in \mathbb{R}^D$ is a $D$-dimensional signal or a vectorized image\footnote{In this manuscript, we work on 2D images.}, $\filter_k \in \mathbb{R}^M$ is the $k$-th dictionary, $\code_k\in \mathbb{R}^D$ is the sparse code associated with that dictionary,  $\lambda>0$ is a sparsity inducing penalty parameter, $K$ is the number of dictionary filters, and $*$ is the convolution operator. The above model will be applied to all the training images.

The modern approaches exploit Parseval's theorem and introduce two slack variables to separate the non-smooth $L_1$ penalty term and the $L_2$ constraints, making it feasible to be efficiently computed in frequency domain, and to apply splitting strategy to formulate the problem into ADMM framework, where the two subproblems (updating $\code$ and updating $\filter$) are jointly solved by coordinate descent~\cite{bristow2013fast,heide2015fast,wohlberg2016efficient}. The model and existing frequency solvers suffer from several issues:

\begin{itemize}
  \item While CSC overcomes the independence assumption held in patch-based learning algorithms, far more variables ($K$ times more) are introduced to represent a single image to compensate for this. This creates more severe memory and computational burdens.

  \item We observe through experiments that the vast majority of the entries of the reconstructed sparse codes do not provide useful information about the represented image. For $K=100$, 99.5\% entries are not informative. This indicates that the subproblem for updating $\code$ solves a highly sparse LASSO problem. Transforming the problem into frequency domain imposes restriction on the use of optimization strategies for this very specific problem.

  \item While prior work shows its efficiency in solving the CSC problem in the frequency domain, this is only applicable for updating $\code$, and does not hold for updating $\filter$. The dictionary filters usually have much smaller spatial support than the dimension size of the sparse codes ($M \ll D$). However, in order to tackle the problem in the frequency domain, it requires to process the $\filter$-subproblem in the same dimensions as the sparse codes, and then project the results onto its small spatial support.
\end{itemize}

### 2018-05-15 Stochastic Convolutional Sparse Coding/Results.tex
update

\section{Results} \label{sec:result}
\subsection{Sub-Sampling Strategy}
We first evaluate the convergence and effectiveness of the proposed sub-sampling strategy, and compare it with state-of-the-art frequency solver~\cite{heide2015fast}. The proposed algorithm, with different sub-sampling factors, was tested on the fruit dataset~\cite{zeiler2010deconvolutional} which contains 10 training images in the size of $100 \times 100$. The dictionary dimension size is set to $11 \times 11 \times 100$.

{\bfseries Convergence.} Comparisons of the convergence between the proposed method and the state-of-the-art are shown in Fig.~\ref{fig:subsampleResult}. A different selection of the subsampling probability reveals that the proposed strategy will slightly influence the convergence and the training objective of the minimization problem. Specifically, the more subsampled, the relatively slower convergence and the higher objective will be obtained. On the other hand, small subsampling rate will significantly accelerate the computation process, where $10\%$ subsampling achieves about $6 \times$ speedup over no subsampled solver and $2 \times$ speedup over state-of-the-art frequency solvers for one iteration. We observe that a subsampling rate between $0.1$ and $0.2$ delivers empirically good enough convergence in our settings, as well as achieving at least $3 \times$ speedup. It should be clarified that for a better representation, the cropped objectives are plotted after the first iteration, while all of them have the same initial objective. Notice that the comparable method has a relatively high objective at first $6-8$ iterations due to different splitting strategies are applied and subproblems are interleaved on the primary variables~\cite{wohlberg2016efficient}, though it converges to an optimum with comparable objective after 10 iterations. In summary, comparing to the state-of-the-art frequency solver, the proposed method with a subsampling rate of $10\%$ not only spends two times less processing time per iteration, but also needs fewer iterations to converge.

@@

\caption{Left: Convergence comparison between the-state-of-art method\cite{heide2015fast} and the proposed method with different subsampling probability. Right: Learned filters by the proposed method with $p=0.1$ and the comparable method. In these represented learned filters, our method learns more Gabor-like and less noise-like filters. Beyond this, it runs faster with regard to each iteration and shows better convergence behavior.}
\label{fig:subsampleResult}
\end{figure}

@@

\begin{figure}[h]
    \includegraphics[width=1\textwidth]{figure/reconImage.pdf}
    \\
    \resizebox{1\linewidth}{!}{
        \begin{tabular}{|c||c|c|c|c|c|c|c|c|c|c|}
            \cline{1-11}
            Image & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8 & 9 & 10 \\
            \hline
            PSNR~\cite{heide2015fast} & 25.77 & 24.67 & 24.89 & 23.88 & 24.68 & 25.26  & 24.15 & 25.75 & 22.97 & 26.16 \\
            \hline
            PSNR ours & \textbf{25.91} & \textbf{24.72} & \textbf{25.21} & \textbf{24.09} & \textbf{25.00} & \textbf{25.36} & \textbf{24.32} & \textbf{25.80} & \textbf{23.24} & \textbf{26.37} \\
            \hline
        \end{tabular} }
    \caption{Numerical comparisons of the reconstruction quality obtained from the presented filters and its comparison. The reconstructions are performed on $20\%$ randomly observed images, with $\lambda = 0.4$ and $50$ ADMM iterations for both cases. Obtained PSNR values are averaged on 5 trials. Note that none of the testing images are in the training sets.} \label{fig:PSNRrecon}
\end{figure}

@@

{\bfseries Convergence.} Unlike the batch-based learning approaches, the training loss objective of the entire dataset cannot be evaluated since each training step only draws one or a few of the data samples. Otherwise, we can evaluate the learning process on testing images. The loss objectives on testing dataset of the proposed algorithm and prior published online CSC model solved in frequency domain~\citep{liu-2018-first}, with respect to iterations, are shown in Fig.~\ref{fig:onlineSmall}. In the same figure, we also keep track of the capability of the updated filters in the learning process to sparsely represent the testing images, which are demonstrated by its PSNR. These two approaches stop at optimum positions with close testing objective values, though, they exhibit slightly different convergence behaviors with respect to iterations. The final testing PSNR values for both methods also reveal the similar reconstruction performance of the learned filters. As for the runtime comparison, the proposed method runs at least 5 times faster than its comparative.

@@

\caption{Left: Convergence of the test set objectives for the proposed method and its comparison~\cite{liu-2018-first}. Right: Testing PSNR with respect to execution time. Our method has the similar outcomes as the comparable method, while it achieves $5 \times$ speedup. }
\label{fig:onlineSmall}
\end{figure}

@@

\begin{subfigure}{1\textwidth}
    \centering
  \includegraphics[width=0.6\linewidth]{figure/online100-filter.pdf}
\end{subfigure}
\end{minipage}
\begin{minipage}{0.5\textwidth}
\centering
\includegraphics[width=1\linewidth]{figure/online400-filter.pdf}
\end{minipage}

\caption{Filters learned from 1000 images by the proposed method and its comparison. (a) The under-complete dictionary ($11 \times 11 \times 100$) learned by~\cite{liu-2018-first}; (b) The under-complete dictionary ($11 \times 11 \times 100$) learned by the proposed method. (c) The over-complete dictionary ($11 \times 11 \times 400$) learned by the proposed method. These under-complete dictionaries, mainly composed of gabor-like filters, can be seen as a subset of the represented over-complete dictionary, which also contains a number of low contrast image features.}
\label{fig:overCompleteDic}
\end{figure}

We demonstrate its capability on 1000 image patches with the size of $100 \times 100$, and learn $11 \times 11 \times 400$ over-complete dictionary, all of which are shown in Fig.~\ref{fig:overCompleteDic} . As a comparison, 100 learned filters by the same algorithm and its comparable approach are also represented for a visual comparison. As can be observed, both of the approaches learn visually similar under-complete dictionaries, while the proposed method shows a $5 \times$ speedup over the other. Furthermore, the learned over-complete dictionary is composed of the gabor-like image features as represented in the under-complete dictionaries, as well as a number of low contrast features which are rarely existed in those under-complete dictionaries. These additional feature information would play an essential role to better represent the natural images, which can be demonstrated in Fig.~\ref{fig:overCompleteDic}.

There is a bottleneck for under-complete dictionary revealed in the left hand side of fig.~\ref{fig:overComDicAndMinibatch}, that no more apparent progress could be observed when the number of training images is higher than a specific value for both of the online approaches. However, owing to more abundant filters, learning over-complete dictionary overcomes this bottleneck, and it shows a considerable improvement with regard to PSNR to have a sparse representation of the testing images.

{\bfseries Mini-batch.} In practice, the mini-batch strategy would be preferred for a gain of convergence speed, rather than drawing a single sample in each training step. This is also a standard extension to stochastic algorithms. We denote the mini-batch size as $\eta$. In the proposed algorithm, the time complexity for one step dictionary update will not increase linearly with the increase of $\eta$. Concretely speaking, updating $\code$ is implemented by Cholesky decomposition, and one computation of the matrix factorization can be applied to all of the currently selected batches. Herein, the complexity for updating $\eta$ $\code$s at once is cheaper than $\eta$ times the complexity of updating one $\code$. In addition, the time complexity for updating $\filter$ is not affected by the value of $\eta$, which will be executed only once in each training step regardless of $\eta$. One exception is to update surrogate matrices, the complexity of which is $\eta$ linearly dependent, while this is not dominant in the runtime. The results for various mini-batch sizes are shown in Fig.~\ref{fig:overComDicAndMinibatch}. The learned filters are examined on the test set every $2^i$ iterations and the last iteration, where $i=0,1,...$. Note that larger $\eta$ will result in less number of iterations to process all 1000 images. A larger mini-batch size shows a greater progress for first few training steps, though it takes additional running time and memory. Overall, mini-batched updates provides a more runtime efficient learning progress in online-based CSC model, and $\eta=20$ achieves one magnitude speedup to reach a condition with same level convergence.

@@

>>>>

\section{Results} \label{sec:result}
\subsection{Subsampling Strategy}
We first evaluate the convergence and effectiveness of the proposed subsampling strategy, and compare it with the state-of-the-art frequency solver~\cite{heide2015fast}. The proposed algorithm, with different sub-sampling factors, was tested on the fruit dataset~\cite{zeiler2010deconvolutional} which contains 10 training images in the size of $100 \times 100$. The dictionary dimension size is set to $11 \times 11 \times 100$.

{\bfseries Convergence.} Comparisons of the convergence between the proposed method and the state-of-the-art are shown in Fig.~\ref{fig:subsampleResult}. A different selection of the subsampling probability reveals that the proposed strategy will slightly influence the convergence and the training objective of the minimization problem. Specifically, the more subsampled, the relatively slower convergence and the higher objective will be obtained. On the other hand, small subsampling rate will significantly accelerate the computation process, where $10\%$ subsampling achieves about $6 \times$ speedup over no subsampled solver and $2 \times$ speedup over state-of-the-art frequency solver for one iteration. We observe that a subsampling rate between $0.1$ and $0.2$ delivers empirically good enough convergence in our settings, as well as achieving at least $3 \times$ speedup. In general, the proposed method converges within 10 iterations, while the comparable method requires roughly 13 iterations to reach converge. Notice that the comparable method has a relatively high objective at first $6-8$ iterations due to different splitting strategies are applied and subproblems are interleaved on the primary variables~\cite{wohlberg2016efficient}, though it converges to an optimum with comparable objective after 10 iterations. In summary, comparing to the state-of-the-art frequency solver, the proposed method with a subsampling rate of $10\%$ not only spends two times less processing time per iteration, but also needs fewer iterations to converge.

@@

\caption{Left: Convergence comparison between the-state-of-art method~\cite{heide2015fast} and the proposed method with different subsampling probability. Right: Learned filters by the proposed method with $p=0.1$ and the comparable method. In these represented learned filters, our method learns more Gabor-like and less noise-like filters. Beyond this, it runs faster with regard to each iteration and shows better convergence behavior.}
\label{fig:subsampleResult}
\end{figure}

@@

\begin{figure}[h]
    \includegraphics[width=1\textwidth]{figure/reconImage.pdf}
    \\
    \\
    \resizebox{1\linewidth}{!}{
        \begin{tabular}{|c||c|c|c|c|c|c|c|c|c|c|}
            \cline{1-11}
            Image & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8 & 9 & 10 \\
            \hline
            PSNR~\cite{heide2015fast} & 29.54 & 28.16 & 29.40 & 29.22 & 28.89 & 29.29 & 28.10  & 29.57 & 27.46 & 30.72 \\
            \hline
            PSNR ours & \textbf{29.65} & \textbf{28.18} & \textbf{29.52} & \textbf{29.43} & \textbf{28.99} & \textbf{29.38} & \textbf{28.19} & \textbf{29.63} & \textbf{27.62} & \textbf{30.98} \\
            \hline
        \end{tabular} }
    \caption{Numerical comparisons of the reconstruction quality obtained from the presented filters and its comparison. The reconstructions are performed on $50\%$ randomly observed images, with $\lambda = 0.4$ and $50$ ADMM iterations for both cases. Obtained PSNR values are averaged on 5 trials. Note that none of the testing images are in the training sets.} \label{fig:PSNRrecon}
\end{figure}

@@

{\bfseries Convergence.} Unlike the batch-based learning approaches, the training loss objective of the entire dataset cannot be evaluated since each training step only draws one or a few of the data samples. Nevertheless, we can evaluate the learning process on testing images. The loss objectives on testing dataset of the proposed online learning algorithm and prior published online CSC model solved in frequency domain~\citep{liu-2018-first}, with respect to iterations, are shown in Fig.~\ref{fig:onlineSmall}. In the same figure, we also keep track of the capability of the updated filters in the learning process to sparsely represent the testing images, which are demonstrated by its PSNR with respect to time. These two approaches stop at optimum positions with close testing objective values, though, they exhibit slightly different convergence behaviors with respect to iterations. The final testing PSNR values for both methods also reveal the similar reconstruction performance of the learned filters. As for the runtime comparison, the proposed method runs at least 5 times faster than its comparative.

@@

\caption{Left: Convergence of the test set objectives for our method (SOCSC) and the state-of-the-art approach~\cite{liu-2018-first}. Right: Testing PSNR with respect to execution time. While the quality of the output is comparable, our method achieves $5 \times$ speedup. \peter{Are the colors correct? The other method seems to be better!}}
\label{fig:onlineSmall}
\end{figure}

@@

\begin{subfigure}{1\textwidth}
    \centering
  \includegraphics[width=0.6\linewidth]{figure/online100-filter.pdf}
\end{subfigure}
\end{minipage}
\begin{minipage}{0.5\textwidth}
\centering
\includegraphics[width=1\linewidth]{figure/online400-filter.pdf}
\end{minipage}
    \\
    \\
    \\
    \resizebox{1\linewidth}{!}{
        \begin{tabular}{|c||c|c|c|c|c|c|c|c|c|c|}
            \cline{1-11}
            Image & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8 & 9 & 10 \\
            \hline
            PSNR by (a) & 29.58 & 28.19 & 29.44 & 29.63 & 28.89 & 29.33 & 28.13  & 30.14 & 27.42 & 30.89 \\
            \hline
            PSNR by (b) & 29.63 & 28.22 & 29.57 & 29.90 & 29.12 & 29.59 & 28.05  & 30.17 & 27.53 & 31.08 \\
            \hline
            PSNR by (c) & \textbf{30.24} & \textbf{28.34} & \textbf{29.95}  & \textbf{30.30} & \textbf{29.43} & \textbf{29.96} & \textbf{28.24} & \textbf{30.57} & \textbf{27.72} & \textbf{31.67} \\
            \hline
        \end{tabular} }

\caption{Filters learned from 1000 images by our method and the state of-the-art method~\cite{liu-2018-first}, and their corresponding reconstruction quality. (a) The under-complete dictionary ($11 \times 11 \times 100$) learned by~\cite{liu-2018-first}; (b) The under-complete dictionary ($11 \times 11 \times 100$) learned by the proposed method. (c) The over-complete dictionary ($11 \times 11 \times 400$) learned by the proposed method. These under-complete dictionaries, mainly composed of Gabor-like filters, can be seen as a subset of the represented over-complete dictionary, which also contains a number of low contrast image features.
\peter{Include these filters in the supplementary, but make them much bigger so that people can see the deltails.}
}
\label{fig:overCompleteDic}
\end{figure}

We demonstrate its capability on 1000 image patches with the size of $100 \times 100$, and learn $11 \times 11 \times 400$ over-complete dictionary, all of which are shown in Fig.~\ref{fig:overCompleteDic} . For a visual comparison, 100 learned filters by the same algorithm and its comparable approach are also represented for a visual comparison. As can be observed, both of the approaches learn visually similar under-complete dictionary, while the proposed method shows a $5 \times$ speedup over the competitive one. Furthermore, the learned over-complete dictionary is composed of the Gabor-like image features as represented in the under-complete dictionaries, as well as a number of low contrast features which are rarely existed in those under-complete dictionaries. These additional feature information would play an essential role to better represent the natural images, which can be demonstrated by the reconstruction quality shown in Fig.~\ref{fig:overCompleteDic}.

We further show a bottleneck revealed by the under-complete dictionary in the online-based CSC model. The left hand side of Fig.~\ref{fig:overComDicAndMinibatch} demonstrates that no more apparent progress could be observed when the number of training images is higher than a specific value for both of the online approaches. However, owing to more abundant filters, learning over-complete dictionary overcomes this bottleneck, and it shows a considerable improvement with regard to PSNR to have a sparse representation of the testing images during the learning process. This is also verified by applying these learned dictionaries to the image inpainting applications as discussed above.

@@

{\bfseries Mini-batching.} In practice, a mini-batching strategy would be preferred in order to gain advantage from modern parallel computing architectures. This is also a standard extension to stochastic optimization algorithms \cite{Takac2013, PCDM, SCSG}. We denote the mini-batch size as $\eta$. In the proposed algorithm, the time complexity for one step dictionary update will not increase linearly with the increase of $\eta$. Concretely speaking, updating $\code$ is implemented by Cholesky decomposition, and one computation of the matrix factorization can be applied to all of the currently selected batches. Herein, the complexity for doing updating $\code$ $\eta$ times all together is cheaper than $\eta$ times the complexity of updating one $\code$. In addition, the time complexity for updating $\filter$ is not affected by the value of $\eta$, which will be executed only once in each training step regardless of $\eta$. One exception is to update surrogate matrices, the complexity of which is $\eta$ linearly dependent, while this is not dominant in the runtime. The learned filters are examined on the test sets every $2^i$ iterations and the last iteration, where $i=0,1,...$. Note that larger $\eta$ will result in less number of iterations to process all 1000 images. The results for various mini-batch sizes are shown in Fig.~\ref{fig:overComDicAndMinibatch}. A larger mini-batch size shows a greater progress for first few training steps, though it takes additional running time and memory. Overall, mini-batched updates provides a more runtime efficient learning progress in online-based CSC model, and in the test examples, $\eta=20$ achieves one magnitude speedup over $\eta=1$ to reach a condition with the same level convergence.

### 2018-05-15 Stochastic Convolutional Sparse Coding/SCSCmodel.tex
update

\section{Stochastic Convolutional Sparse Coding}
\subsection{The Model}
Since the images are represented by a set of highly sparse codes, we have observed that they can be well reconstructed by a random subset of the sparse codes. Based on this observation, we propose a variant of the $\code$-subproblem using subsampling strategy, and formulate the following modified minimization problem:
\begin{equation} \label{eq:updatingCode}
    \minimize{\code} \frac{1}{2}\|\signal - (\Filter \mask^T)(\mask \code) \|_2^2 + \lambda \|\mask \code\|_1,
\end{equation}
where $\Filter = [\Filter_1 ... \Filter_K] \in \mathbb{R}^{D \times DK}$, $\code = [\code_1^T ... \code_K^T]^T \in \mathbb{R}^{DK \times 1}$, the convolution operator are formulated in matrix multiplication so that $ \Filter \code = \sum_{k=1}^{K} \filter_k * \code_k$.  $\mask$ is a randomized subsampling diagonal matrix with $1$ on the entries corresponding to the sampled sparse codes, performing the function of choosing a random portion of the sparse codes for reconstruction. Herein, the $\code$-subproblem only computes the sparse codes in the chosen positions. Due to the introduced subsampling matrix, the convolution operator will not hold, hence it cannot be solved in the frequency domain. This subsampling strategy is akin to the dropout technique~\cite{srivastava2014dropout}, which has been widely used to prevent overfitting when training deep neural networks. However, different from dropout that generally needs more time to train a neural network than standard approaches, with a good selection of subsampling probability $p$, the proposed method shows less per iteration running time and faster convergence than state-of-the-arts (see section~\ref{sec:result}).

After obtaining the subsampled sparse codes, we can then project them onto its original spatial support with $0$ filling. Afterwards, the dictionary can be updated by:
\begin{equation} \label{eq:updatingFilter}
    \minimize{\filter} \frac{1}{2}\|\signal - \Code \filter \|_2^2, ~\subjectto ~ \|\filter_k\|^2_2 \leq 1 ~ \forall k \in {1,...K},
\end{equation}
where $\Code= [\Code_1 ... \Code_K] \in \mathbb{R}^{D \times MK}$, constructed from computed $\code$, and $\filter= [\filter_1^T ... \filter_K^T]^T \in \mathbb{R}^{MK \times 1}$ such that $ \Code \filter = \sum_{k=1}^{K} \filter_k * \code_k$.

Like the approach in~\cite{heide2015fast}, we interleave on the above two subproblems to tackle the bi-convex problem, and each of them will be solved by ADMM. The joint optimization framework will be implemented in the fashions of batch mode and online mode.

@@

\subsection{Stochastic Batch CSC (SBCSC)}
We first formulate the batch-mode of the proposed method as shown in Algorithm~\ref{algo:SBCSC}, where $N$ is the number of total input images, $\code^i$ is the sparse code associated with $i$-th image, and $p$ is the uniform probability for one code been selected. Every outer loop involves all of the training images, thus it would be expensive in processing on these simultaneously and also runs out of memory quickly with the increase of the number of images. The batch-based learning algorithm lacks of the capability to scale up to very large datasets or to handle dynamically changed training data.

\subsection{Stochastic Online CSC (SOCSC)}
We can further tackle the proposed problem in online fashion for a gain of scalability. In the online learning setting, each iteration only draws one or a subset (mini-batch) of the total training images, hence the complexity per loop is independent of the training sample size. Then, given the sampled image $\signal^t$ at $t$-th iteration, we can compute the corresponding sparse codes $\code^t$ by
\begin{equation} \label{eq:updatingCodeOnline}
    \code^t = \minimize{\code} \frac{1}{2}\|\signal^t - (\Filter^{t-1} \mask^T)(\mask \code) \|_2^2 + \lambda \|\mask \code\|_1,
\end{equation}
where $\Filter^{t-1}$ is constructed from the dictionary learned by $(t-1)$th iteration. After obtaining the sparse codes, the dictionary is updated by:
\begin{equation}
    \filter^t = \minimize{\filter} \frac{1}{2t}\sum_{i=1}^{t} \|\signal^i - \Code^i \filter \|_2^2, ~\subjectto ~ \|\filter_k\|^2_2 \leq 1 ~ \forall k \in {1,...K}.
\end{equation}
Notice that updating dictionary involves all of the past training images and sparse codes. As shown in~\cite{mairal2009online,mairal2010online}, we can get rid of explicitly storing those data by introducing two surrogate matrices $\surC \in \mathbb{R}^{KM \times KM}$ and $\surB \in \mathbb{R}^{KM \times 1}$, which carry all of the required information for updating $\filter$, and can be iteratively updated by:
\begin{equation} \label{eq:updateSur}
    \surC^t  = \frac{t-1}{t} \surC^{t-1} + \frac{1}{t}(\Code^t)^T\Code^t ~~~~~~~~~ \surB^t  = \frac{t-1}{t} \surB^{t-1} + \frac{1}{t}(\Code^t)^T\signal^t
\end{equation}
If so, the updated dictionary can be obtained by solving:
\begin{equation} \label{eq:updatingFilterOnline}
    \filter^t = \minimize{\filter} \frac{1}{2} \filter^T\surC\filter - \filter^T\surB, ~ \subjectto ~ \|\filter_k\|_2^2 \leq 1 ~ \forall k \in {1,...K}
\end{equation}

@@

All training and evaluation processes in this manuscript are performed on contrast normalized images. %We also tried to learning filters from images without pre-processing with little parameters tuning, which successfully learns similar Gabor-like features as learned from contrast normalized images, as well as low frequency filters. The results regarding to this will not be discussed in this work, while it can be applied to other applications where low frequency parts also need to be taken into account. %
It should be noted that matrix $\Code$ needs to be explicitly constructed and it may lead to a considerable computational burden, resulting in a heavy overhead if processed in an inefficient way. In our implementation, the vector-to-matrix mapping information is cached for an efficient indexing. As for matrix $\Filter$, we only explicitly construct matrix $(\Filter \mask^T)$ rather than $\Filter$ and $\mask$ separately for the sake of efficiency, and the similar indexing strategy is applied.

>>>>

\section{Stochastic Convolutional Sparse Coding}
\subsection{The Model}
Since the images are represented by a set of highly sparse codes, we have observed that they can be well reconstructed by a random subset of the sparse codes. Based on this observation, we propose a variant of the $\code$-subproblem using subsampling strategy, and formulate the following modified minimization problem:
\begin{equation} \label{eq:updatingCode}
    \minimize{\code} \frac{1}{2}\|\signal - (\Filter \mask^\top)(\mask \code) \|_2^2 + \lambda \|\mask \code\|_1,
\end{equation}
where $\Filter = [\Filter_1, \dots, \Filter_K] \in \mathbb{R}^{D \times DK}$, $\code = [\code_1, \dots, \code_K] \in \mathbb{R}^{DK}$, the convolution operator are formulated in matrix multiplication so that $ \Filter \code = \sum_{k=1}^{K} \filter_k * \code_k$.  Matrix $\mask$ is a random diagonal subsampling  matrix with ones on the entries corresponding to the sampled sparse codes and zeros elswhere, performing the function of choosing a random portion of the sparse codes for reconstruction. Herein, the $\code$-subproblem only computes the sparse codes in the chosen positions. Due to the introduced subsampling matrix, the convolution operator will not hold, hence it cannot be solved in the frequency domain. This stochastic subsampling strategy is akin to the dropout technique~\cite{srivastava2014dropout}, which has been widely used to prevent overfitting when training deep neural networks. However, different from dropout that generally needs more time to train a neural network than standard approaches, with a good selection of subsampling probability $p$, the proposed method shows less per iteration running time and faster convergence than state-of-the-arts (see Section~\ref{sec:result}).

After obtaining the subsampled sparse codes, we can then project them onto its original spatial support with $0$ filling. Afterwards, the dictionary can be updated by:
\begin{equation} \label{eq:updatingFilter}
    \minimize{\filter} \frac{1}{2}\|\signal - \Code \filter \|_2^2 \quad \text{subject to} \quad \|\filter_k\|^2_2 \leq 1 ~ \forall k \in \{1,\dots,K\},
\end{equation}
where $\Code= [\Code_1, \dots, \Code_K] \in \mathbb{R}^{D \times MK}$, constructed from computed $\code$, and $\filter= [\filter_1,\dots,\filter_K] \in \mathbb{R}^{MK}$ such that $ \Code \filter = \sum_{k=1}^{K} \filter_k * \code_k$.

Similarly to the approach in~\cite{heide2015fast}, we alternate the above two subproblems to tackle the bi-convex problem, and each of them will be solved by ADMM. The joint optimization framework will be implemented in the fashions of {\em batch mode} (stochastic batch CSC) and {\em online mode} (stochastic online CSC).

@@

\subsection{Stochastic Batch CSC}
We first formulate the batch-mode of the proposed method as shown in Algorithm~\ref{algo:SBCSC}, where $N$ is the number of total input images, $\code^i$ is the sparse code associated with $i$-th image, and $p$ is the uniform probability for one code been selected. Every outer loop involves all of the training images, thus it would be expensive in processing on these simultaneously and also runs out of memory quickly with the increase of the number of images. The batch-based learning algorithm lacks of the capability to scale up to very large datasets or to handle dynamically changed training data.

\subsection{Stochastic Online CSC}
We can further tackle the proposed problem in online fashion for a gain of scalability. In the online learning setting, each iteration only draws one or a subset (mini-batch) of the total training images, hence the complexity per loop is independent of the training sample size. Then, given the sampled image $\signal^t$ at $t$-th iteration, we can compute the corresponding sparse codes $\code^t$ by
\begin{equation} \label{eq:updatingCodeOnline}
    \code^t = \minimize{\code} \frac{1}{2}\|\signal^t - (\Filter^{t-1} \mask^\top)(\mask \code) \|_2^2 + \lambda \|\mask \code\|_1,
\end{equation}
where $\Filter^{t-1}$ is constructed from the dictionary learned by $(t-1)$th iteration. After obtaining the sparse codes, the dictionary is updated by:
\begin{equation}
    \filter^t = \minimize{\filter} \frac{1}{2t}\sum_{i=1}^{t} \|\signal^i - \Code^i \filter \|_2^2 \quad \text{subject to} \quad \|\filter_k\|^2_2 \leq 1 ~ \forall k \in \{1,\dots,K\}.
\end{equation}
Notice that updating dictionary involves all of the past training images and sparse codes. As shown in~\cite{mairal2009online,mairal2010online}, we can get rid of explicitly storing those data by introducing two surrogate matrices $\surC \in \mathbb{R}^{KM \times KM}$ and $\surB \in \mathbb{R}^{KM \times 1}$, which carry all of the required information for updating $\filter$, and can be iteratively updated by:
\begin{equation} \label{eq:updateSur}
    \surC^t  = \frac{t-1}{t} \surC^{t-1} + \frac{1}{t}(\Code^t)^\top \Code^t ~~~~~~~~~ \surB^t  = \frac{t-1}{t} \surB^{t-1} + \frac{1}{t}(\Code^t)^\top \signal^t
\end{equation}
If so, the updated dictionary can be obtained by solving:
\begin{equation} \label{eq:updatingFilterOnline}
    \filter^t = \minimize{\filter} \frac{1}{2} \filter^t \surC\filter - \filter^t\surB \quad \text{subject to} \quad \|\filter_k\|_2^2 \leq 1 ~ \forall k \in \{1,\dots,K\}.
\end{equation}

\peter{The above equations has issues.}

@@

All training and evaluation processes in this manuscript are performed on contrast normalized images. %We also tried to learning filters from images without pre-processing with little parameters tuning, which successfully learns similar Gabor-like features as learned from contrast normalized images, as well as low frequency filters. The results regarding to this will not be discussed in this work, while it can be applied to other applications where low frequency parts also need to be taken into account. %
It should be noted that matrix $\Code$ needs to be explicitly constructed and it may lead to a considerable computational burden, resulting in a heavy overhead if processed in an inefficient way. In our implementation, the vector-to-matrix mapping information is cached for an efficient indexing. As for matrix $\Filter$, we only explicitly construct matrix $(\Filter \mask^\top)$ rather than $\Filter$ and $\mask$ separately for the sake of efficiency, and the similar indexing strategy is applied.

### 2018-05-15 nips2018/nips_2018.tex
update

@@

\begin{abstract}

\end{abstract}

@@

% \input{Discussions}

@@

>>>>

% Commenting
\usepackage[colorinlistoftodos,bordercolor=orange,backgroundcolor=orange!20,linecolor=orange,textsize=scriptsize]{todonotes}
\newcommand{\peter}[1]{\todo[inline]{\textbf{Peter: }#1}} 
\newcommand{\yinhui}[1]{\todo[inline]{\textbf{Yu: }#1}}

@@

% snr onlineVSbatch online400Small

\begin{abstract}
The minimization problems formulated in Convolutional Sparse Coding (CSC) have been solved in frequency domain by modern approaches owing to the improved computing efficiency, while it automatically assumes circular boundary conditions. In this work, we propose a novel stochastic spatial-domain solver where a randomized subsampling strategy is introduced in the subproblem of computing sparse codes. The proposed method experimentally shows a faster convergence rate compared to the state-of-the-art, and with a reasonable selection for the subsampling probability, it also achieves a better runtime efficiency. Moreover, we extend the proposed strategy in conjunction with online learning, scaling the CSC model up to arbitrary sample sizes. This combination leads to a significant speedup over the online CSC algorithm solved in frequency domain. We then learn the over-complete dictionary from thousand of images, which demonstrates a better sparse representation of the natural images on account of more abundant filters.
\end{abstract}

@@

\input{Conclusion}

@@

\clearpage
\appendix
\section{More Experiments}

\peter{Include more experiments here}
